BOLD Parameter Estimation Using Particle Filters

### 2010-07-21 doc/thesis/methods.tex
corrected the intro, and moved the noise description out to methods

\chapter{Methods}
Although the particle filter I used for this work is a standard Regularized
Particle filter, as described in \cite{Arulampalam2002a}, optimizing the 
particle filter for use with FMRI data is non-trivial.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c || c | c | c |}
\hline 
Parameter & Distribution & $\mu$ & $\sigma$ \\
\hline
$\tau_0$ & Gamma & .98 & .25 \\
$\alpha$ & Gamma & .33 & .045\\
$E_0$    & Gamma & .34 & .03  \\
$V_0$    & Gamma & .04 & .03 \\
$\tau_s$ & Gamma & 1.54  & .25\\
$\tau_f$ & Gamma & 2.46  & .25\\
$\epsilon$ & Gamma & .7  & .6 \\
\hline
\end{tabular}
\caption{}
\label{tab:Prior} 
\end{table}

@@

After trial and error, I chose the weight standard deviation to be $.005$. Obviously
the case of the Cauchy, this is simply the scale parameter. While I 
did attempt to automatically set the standard deviation, I found that the algorithm
became too unpredictable. Essentially, if the weight function isn't
fixed across voxels, very noisy time series with no actual signal will 
converge to nonsensical results. 
In the future, it may be possible to set the standard deviation by
taking a small sample from "resting" data and using the sample standard deviation.
Since this is the first attempt at using particle filters for modeling the 
BOLD model, in this work I set the standard deviation manually at $.005$,
because it gives a more consistency and control.

\subsection{Classical De-trending}
\label{sec:Detrend}
The non-stationary
aspect of a Weiner process, presumably the result of integrating some
$\nu_x$ is difficult to compensate for, and so various methods
have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
of FMRI. In some studies, as many as half the voxels benefit from detrending, presenting
a serious barrier to inference (\cite{Smith2007}). All the existing methods are performed
during the preprocessing stage, rather than as an integral part of analyzing the BOLD
signal. There is no shortage of theories on the "best" method of detrending; however
in a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
a spline works the best. The benefit of the spline versus wavelets, high pass 
filtering or other DC removal techniques is that the frequency response is not set.
Rather, the spline is adaptive to the input, having a low cut off if the signal's 
median stays constant but a high cut off frequency if the signal's median tends to
shift heavily over the course of time. In spite of this, the spline will still remove some
amount of signal, just like all of these methods, which is why I considered the method proposed
in \autoref{sec:Methods Delta Based Inference}.

The method I used to calculate the spline was picking one knot for every 15
measurements in an image. Thus a 10 minute session at a repetition time of 
2.1 seconds would have 19 knots. The knot first and last knots were each 
given half the number of samples as the rest of the knots; which were all 
located at the center of their sample group. The median of each sample group
was then taken and used as the magnitude for the group. Taking the median 
versus the mean seemed to work better, given the presence of outliers. It seems
that there is potential to optimize the spline further using a canonical 
HRF to find resting points; however, for this to work the experiment would have
to be designed with this in mind.

When using the median-based spline techniques, the normalized signal will,
by definition have a median near zero. 
The problem with this is this is not the "natural" BOLD signal. More specifically,
when the signal is inactive, the BOLD response \emph{should} be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
After removing the spline, the BOLD resting state will be below 0\%.
This is problematic because it reduces the ability of an algorithm to learn.
One quick and dirty solution is to add an arbitrary constant to each BOLD response. 
Of course this won't scale to whole-brain analysis, so a more effective technique 
is adding a DC gain parameter to the BOLD model. 
Like all the other model parameters, given enough measurements, correct
value may be found. On the downside, adding another dimension increases the
complexity of the model, for a variable that is relatively obvious by direct
observation.

@@

\begin{figure}
\caption{image of de-spline'd lines with "true" lines}
\label{fig:DesplineQuality}
\end{figure}

@@

\subsection{Delta Based Inference}
\label{sec:Methods Delta Based Inference}
The alternative to these sorts of low frequency manipulation is to
go around the problem in another way. Here, I propose and test a 
different method of dealing with the so called "drift". 
Instead of comparing the direct output of the particle filter with the direct
measurement, the algorithm would compare the change in signal over a single TR,
with the result of integrating the model for the same period. 
In most signal processing cases this would be foolish, but that is because the 
general assumption is that all noise is high frequency. Considering 
the fact that every BOLD analysis pipeline uses a high pass filter,
whereas low poss temporal filter are rarely applied, it makes sense
that a derivative type method could work. The benefit of particle filters
is that they are extremely robust method of inference, and I would assert 
that the particle filter ought to be given as \emph{raw} data as possible. 
While taking direct measurements
without de-trending would give awful results, using the difference removes the 
DC component and turns what is usually assumed to be a Weiner process into 
a simple Gaussian random variable.

@@

\section{Preprocessing}
\label{sec:Methods Preprocessing}
As discussed in the section on de-trending, the normal pipeline for analyzing
FMRI involves a great deal of preprocessing. The first and most important
task is motion correction. To do this, a single volume in time is chosen, and
volumes at every other time are registered to this one volume. This takes care
of motion by the patient as well as some small changes in the magnetic
fields that cause the image to shift. After this, a structural image of the
patient is often registered to an atlas so that a grey-matter mask may
be created. Since all the "activation" should occur in grey matter voxels,
it made sense to restrict analysis to only these sections; although
there may be partially grey-matter areas that are missed. This significantly
eases computation time though.

In normal statistical parametric mapping, a gaussian smoothing
filter is applied across the image as discussed in \autoref{sec:RFT}.
After this, detrending is performed discussed by \autoref{sec:Detrend}.
Recall that FMRI signal levels are unit-less and even though detrending isn't
always necessary, at the very least the data must be converted 
into \% difference from the baseline. Changing to 
\% difference removes no real information
from signal. This is the signal that was input into the delta based 
particle filter. Of course, most of the time analysis is performed on the
direct signal; which mandates the removal of low frequency drift.
The generally accepted method is to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.

>>>>

\chapter{Methods}
\label{sec:Methods}
Although the particle filter I used for this work is a standard Regularized
Particle filter, as described in \cite{Arulampalam2002a}, optimizing the 
particle filter for use with FMRI data is non-trivial.

@@

\begin{table}[t]
\centering
\begin{tabular}{|c || c | c | c |}
\hline 
Parameter & Distribution & $\mu$ & $\sigma$ \\
\hline
$\tau_0$ & Gamma & .98 & .25 \\
$\alpha$ & Gamma & .33 & .045\\
$E_0$    & Gamma & .34 & .03  \\
$V_0$    & Gamma & .04 & .03 \\
$\tau_s$ & Gamma & 1.54  & .25\\
$\tau_f$ & Gamma & 2.46  & .25\\
$\epsilon$ & Gamma & .7  & .6 \\
\hline
\end{tabular}
\caption{Prior distributions used in the particle filter.}
\label{tab:Prior} 
\end{table}

@@

After trial and error, I chose the weight standard deviation to be $.005$. Obviously
the case of the Cauchy, this is simply the scale parameter. While I 
did attempt to automatically set the standard deviation, I found that the algorithm
became too unpredictable. Essentially, if the weight function isn't
fixed across voxels, very noisy time series with no actual signal will 
converge to nonsensical results. 
In the future, it may be possible to set the standard deviation by
taking a small sample from resting data and using the sample standard deviation.
Since this is the first attempt at using particle filters for modeling the 
BOLD model, in this work I set the standard deviation manually at $.005$,
because it gives a more consistency and control.

\section{BOLD Noise}
\label{sec:Introduction Noise}
As demonstrated in \autoref{sec:BOLD Physiology} the BOLD response has been
extensively studied and despite minor discrepancies, the cause of the BOLD 
signal is well known. However, as FMRI detects an  
aggregate signal over the space of cubic centimeters, there are
plenty of sources of noise. Though local neurons act
together (i.e. around the same time), the density of neurons, the
density of capillaries, and slight differences in activation across 
a particular voxel can all lead to signal attenuation and noise.

A particularly difficult form of noise present in FMRI is a low frequency
drift, often characterized as a Wiener process (\cite{Riera2004}). 
Though not present in all regions, as much as ten to fifteen percent
of voxels can be affected (\cite{Tanabe2002}), thus it is prevalent enough to cause significant
inference problems \cite{Smith2007}. It is still not
clear what exactly causes this noise noise comes from, although it is possible it is 
the result of magnets heating up, or some distortion in magnetic
fields \cite{Smith2007}. It is clear that this drift signal is not solely
due to a physiological effects, given its presence in cadavers and phantoms 
\cite{Smith1999}. Interestingly, it is usually spatially correlated, and
more prevalent at interfaces between regions, though 
by no means limited to such areas. Though one potential source
could be slight movement, given that co-registration of volumes to a single time
point is standard, this seems unlikely. Regardless, the problem mandates
the use of a high pass filter to make inference reasonably powered
\cite{Smith2007}.

In order to characterize the noise, I analyzed resting state data.
During resting state, the patient is shown no images, and he is asked
to avoid movement and complex thought, to the best of his abilities.
Some limitations exist, for instance EPI sequences are extremely loud
which could cause some stimulus in the auditory cortex, and of course
very long sequences could give the patient time for his mind to wander.
Overall though there should be no activation, and thus the signal will
consist entirely of noise. Therefore resting state data is perfect
for analyzing the noise distribution. The locations were chosen from
various points around the brain, all in grey matter voxels. The time
series were also chosen because they were representative of different types
of noise I found in the resting state data.

Because most methods (including the one used in this paper)
assume the noise realizations are independent of each other, the auto-
correlation is of particular interest (which is a necessary but not
sufficient condition for independence). Gaussianity is also a common
assumption made in studies of FMRI data, though that assumption is not
made in this work. Regardless, comparing the distribution to a Gaussian
is informative, so I used Q-Q plots to compare example data with the
Normal distribution. Additionally, in FMRI data the noise is often considered 
to be Wiener \cite{Riera2003}. Recall that a Wiener random process is
characterized by steps that are Gaussian; in other words the difference between
any two samples is Normally distributed. The simulations discussed in 
\autoref{sec:Single Voxel Simulation} make use of this, 
by adding a Wiener random process to the overall signal. To determine
whether the noise is in fact Wiener, the distribution of 
the difference between adjacent measurements was plotted against
a Gaussian. Wiener processes further assume that the steps are independent;
so the steps should have near flat autocorrelation functions.

Finally, removal of the drift is often
performed with some variation of a high pass filter, 
so I checked the noise distribution after applying such a filter (in
this case the subtraction of a spline, see \autoref{sec:Methods Preprocessing}).
Here I wanted to know how effective my high pass filter at the removal of
Wiener noise.

\autoref{fig:QQDC} shows the 
results with a regression line fit to the points on the Q-Q plot.
Recall that in a Quartile-Quartile (Q-Q) plot, if the points plotted on the 
x-axis and the points
on the y-axis come from the same \emph{type} of distribution then all the points will
fall on a line. Differences in the variance will cause the line to have a slope
other than 1, while differences in the expected value will cause the fitted line
to be shifted. In these Q-Q plots, the points are being compared to the standard
Gaussian distribution, so the quality of the line fit determines how closely the points
fit the Gaussian distribution. Note that in \autoref{fig:QQDC} the points have all been 
normalized (changed to percent difference).
\begin{figure}
\centering
\subfigure[]{\label{fig:QQDC:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_29_49_9}}
\subfigure[]{\label{fig:QQDC:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_34_43_24}}
\subfigure[]{\label{fig:QQDC:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_22_38_23}}
\subfigure[]{\label{fig:QQDC:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009_37_29_24}}

%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_19-24-10.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_20-45-18.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_23-47-18.pdf}}
%\subfigure{\includegraphics[trim=6cm 1cm 0 0cm,width=17cm]{images/noise_0009_35-49-9.pdf}}

\caption{Q-Q Plots of normalized resting state data}
\label{fig:QQDC}
\end{figure}

\begin{figure}
\centering
\subfigure[]{\label{fig:QQDDelta:A}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_29_49_9}}
\subfigure[]{\label{fig:QQDDelta:B}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_34_43_24}}
\subfigure[]{\label{fig:QQDDelta:C}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_22_38_23}}
\subfigure[]{\label{fig:QQDDelta:D}\includegraphics[trim=6cm 1cm 6cm 0,width=14cm]{images/noise2_0009d_37_29_24}}
\caption{Q-Q Plots of resting state data, using the BOLD signal changes}
\label{fig:QQDelta}
\end{figure}

A Wiener process should still conform to the normal distribution, with a 
variance proportional to the run-time. Note that \autoref{fig:QQDC:A} and \autoref{fig:QQDC:B}
are relatively well described by a Gaussian process with a small autocorrelation, 
\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D} are not. In particular the tails of \autoref{fig:QQDC:C}
do not seem to fit the Gaussian well. Also note the significant autocorrelation in
\autoref{fig:QQDC:C} and \autoref{fig:QQDC:D}. As expected, the noise is not strictly
Gaussian white noise.  On the other hand, the steps do conform rather
closely to the normal distribution.
As expected, most of the autocorrelation disappears for the step data. Given
that the steps seem to fit the Normal distribution, the low autocorrelation
indicates that the steps could be Independently Distributed. 
Therefore, the assumption that the noise follows a Wiener process seems to
be correct.

\begin{figure}
\centering
\subfigure[]{\label{fig:QQs:A}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_29_49_9}}
\subfigure[]{\label{fig:QQs:B}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_34_43_24}}
\subfigure[]{\label{fig:QQs:C}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_22_38_23}}
\subfigure[]{\label{fig:QQs:D}\includegraphics[trim=6cm 1cm 6cm 0cm,width=14cm]{images/noise2_0009s_37_29_24}}
\caption{Q-Q Plots of resting state data, after the de-trending}
\label{fig:QQSpline}
\end{figure}

De-trending the time-series by subtracting a spline fit to the distribution
removed much of the autocorrelation present in \autoref{fig:QQDC:C} and \autoref{fig:QQDC:D},
though clearly not perfectly. Though the distributions still do not exactly fit
the Normal, \autoref{fig:QQs:D} clearly is much improved compared to \autoref{fig:QQDC:D}.
In all, the de-trending is effectively removing Wiener noise.

\subsection{Detrending}
\label{sec:Detrend}
The non-stationary
aspect of a Weiner process, presumably the result of integrating some
$\nu_x$ is difficult to compensate for, and so various methods
have been developed to compensate for it. \cite{Tanabe2002} and \cite{Smith1999} have
demonstrated that this component is prevalent, and may in fact be an inherent  characteristic
of FMRI. In some studies, as many as half the voxels benefit from detrending, presenting
a serious barrier to inference (\cite{Smith2007}). All the existing methods are performed
during the preprocessing stage, rather than as an integral part of analyzing the BOLD
signal. There is no shortage of detrending methods; however
in a head to head comparison, \cite{Tanabe2002}, showed that in most cases subtracting off
a spline works the best. The benefit of the spline versus wavelets, high pass 
filtering or other DC removal techniques is that the frequency response is not set.
Rather, the spline is adaptive to the input, having a low cut off if the signal's 
median stays constant but a high cut off frequency if the signal's median tends to
shift heavily over the course of time. In spite of this, the spline will still remove some
amount of signal, just like all of these methods, which is why I considered the method proposed
in \autoref{sec:Methods Delta Based Inference}.

The method I used to calculate the spline was picking one knot for every 20
measurements in an image. Thus a 10 minute session at a repetition time of 
2.1 seconds would have 19 knots. The knot first and last knots were each 
given half the number of samples as the rest of the knots; which were all 
located at the center of their sample group. The median of each sample group
was then taken and used as the magnitude for the group. Taking the median 
versus the mean seemed to work better, given the presence of outliers. It seems
that there is potential to optimize the spline further using a canonical 
HRF to find resting points; however, for this to work the experiment would have
to be designed with this in mind.

When using the median-based spline techniques, the normalized signal will,
by definition have a median near zero. 
The problem with this is this is not the natural state of the BOLD signal. More specifically,
when the signal is inactive, the BOLD response should be at 0\% change from
the base level; activation may then increase, or for short periods decrease from this base.
After removing the spline, the BOLD resting state will be below 0\%.
This is problematic because it reduces the ability of an algorithm to learn.
One quick and dirty solution is to add an arbitrary constant to each BOLD response. 
Of course this won't scale to whole-brain analysis, so a more effective technique 
is adding a DC gain parameter to the BOLD model. 
Like all the other model parameters, given enough measurements, correct
value may be found. On the downside, adding another dimension increases the
complexity of the model, for a variable that is relatively obvious by direct
observation.

@@

\begin{figure}
\caption{Image of de-spline'd lines with true lines}
\label{fig:DesplineQuality}
\end{figure}

@@

\subsection{Delta Based Inference}
\label{sec:Methods Delta Based Inference}
The alternative to these sorts of low frequency manipulation is to
go around the problem in another way. Here, I propose and test a 
different method of dealing with the drift. 
Instead of comparing the direct output of the particle filter with the direct
measurement, the algorithm would compare the change in signal over a single TR,
with the result of integrating the model for the same period. 
In most signal processing cases this would be foolish, but that is because the 
general assumption is that all noise is high frequency. Considering 
the fact that every BOLD analysis pipeline uses a high pass filter,
whereas low poss temporal filter are rarely applied, it makes sense
that a derivative type method could work. The benefit of particle filters
is that they are extremely robust method of inference, and I would assert 
that the particle filter ought to be given as \emph{raw} data as possible. 
While taking direct measurements
without de-trending would give awful results, using the difference removes the 
DC component and turns what is usually assumed to be a Weiner process into 
a simple Gaussian random variable.

@@

\section{Preprocessing}
\label{sec:Methods Preprocessing}
As discussed in the section on de-trending, the normal pipeline for analyzing
FMRI involves a several preprocessing steps. The first and most important
task is motion correction. To do this, a single volume in time is chosen, and
volumes at every other time are registered to this one volume. This corrects
for motion by the patient as well as small changes in the magnetic
fields that cause the image to shift. 
In conventional statistical parametric mapping, a gaussian smoothing
filter is applied across the image as discussed in \autoref{sec:RFT}.
After this, detrending is performed as discussed by \autoref{sec:Detrend}.
Recall that FMRI signal levels are unit-less and though detrending is not
always necessary, at the very least the data must be converted 
into \% difference from the baseline. Changing to 
\% difference removes no real information
from signal. This is the signal that was input into the delta based 
particle filter. Of course, most of the time analysis is performed on the
direct signal; which mandates the removal of low frequency drift.
The generally accepted method is to use a high pass filter, although the
cutoff frequency is application dependent and often applied haphazardly.

### 2010-07-21 doc/thesis/particlefilters.tex
More clean up.
corrected the intro, and moved the noise description out to methods

\section{Resampling}
\label{sec:Particle Filter Resampling}
As a consequence 
of the wide prior distribution (required for a proper discretization of a continuous
distribution), there will be many particles with insignificant weights. While this does help
describe the tails of the distribution, it means a lot of computation will be based.
Instead, it would be preferable if most of the computation is spent on the most probable regions.
Ideally the computation time spent on tails would be proportional to the actual size of the
tails. In this case particle locations would match the true posterior and all weights would
be equal.  The case where a large number of the weights have become irrelevantly small
is called "particle degeneracy". In  \cite{Liu1998b}
an ideal calculation of the "effective" number of particles is found based on the 
particles' "true weight". However, given that only an approximation for the true weight 
exists, they also provide a simple heuristic calculation of $N_{eff}$.
\begin{equation}
N_{eff} \approx \frac{\sum_{i=0}^{N_p} w_i}{\sum_{i=0}^{N_p} w_i^2}
\label{eq:neff}
\end{equation}
Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
$N_{eff}$ drops precipitously.  To alleviate this problem
a common technique known as resampling may be applied. The idea of resampling is to 
draw from the approximate posterior, thus generating a replica of the posterior with 
a better support. Therefore, a new set of particles may be drawn from the empirical
distribution as follows:
\begin{equation}
\hat{x}_j \sim \left(\sum_{i=0}^{N_p} w^i_k\delta(x - x^i_k)\right)
\end{equation}

\begin{algorithm}
\caption{Resampling Algorithm}
\begin{algorithmic}
\STATE Calculate total weight, $W = \sum_{i=0}^{N_p} w^i$
\FORALL{$0 < i < N_p$}
    \STATE Draw $V$ from uniform range $[0, W]$
    \STATE $C = W_t$
    \FORALL{$0 < j < N_p$ and $C < V$}
        \STATE $C = C - w^j$
    \ENDFOR
    \STATE Add $[x^j, \frac{1}{N_p}]$ to the new distribution
\ENDFOR
\STATE 
\end{algorithmic}
\label{alg:Resampling}
\end{algorithm}
For infinite particles this new distribution will match the old.
Unfortunately, this isn't the truth in practice: since the support is
still limited to the original particles, the number of \emph{unique} particles can only go down.
This effect, dubbed "particle impoverishment" can result in excessive quantization
errors in the final distribution. However, there is a solution. Instead of sampling from the
discrete distribution, a smoothing kernel is applied, and particles are drawn from
that distribution. Because it is continuous, particle impoverishment
cannot occur. The easiest way to sample from the continuous distribution is to break the 
re-sampling down into two steps. After calculating an estimate of the scale of the original
distribution, algorithm \autoref{alg:Resampling} is performed. Next, a distribution is generated
based on the variance of the original distributions.
Finally, for each particle in the discretely re-sampled distribution, a sample is drawn from 
the smoothing 
distribution and added to the particle.  The process regularization is defined as:

@@

The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
or a genetic algorithm. Versions of $x$ that are "fit" (give good measurements) spawn more children 
nearby which allow for more accurate estimation near points of high likelihood. 
As the variance of the estimated
$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.

>>>>

\section{Resampling}
\label{sec:Particle Filter Resampling}
As a consequence 
of the wide prior distribution (required for a proper discretization of a continuous
distribution), there will be many particles with insignificant weights. While this does help
describe the tails of the distribution, it means a lot of computation will be based.
Instead, it would be preferable if most of the computation is spent on the most probable regions.
Ideally the computation time spent on tails would be proportional to the actual size of the
tails. In this case particle locations would match the true posterior and all weights would
be equal.  The case where a large number of the weights have become irrelevantly small
is called \emph{particle degeneracy}. In  \cite{Liu1998b}
an ideal calculation of the effective number of particles is found based on the 
particles' true weight. However, given that only an approximation for the true weight 
exists, they also provide a simple heuristic calculation of $N_{eff}$.
\begin{equation}
N_{eff} \approx \frac{\sum_{i=0}^{N_p} w_i}{\sum_{i=0}^{N_p} w_i^2}
\label{eq:neff}
\end{equation}
Any quick run of a particle filter will reveal that unless the prior is particularly accurate,
$N_{eff}$ drops precipitously.  To alleviate this problem
a common technique known as resampling may be applied. The idea of resampling is to 
draw from the approximate posterior, thus generating a replica of the posterior with 
a better support. Therefore, a new set of particles may be drawn from the empirical
distribution as follows:
\begin{equation}
\hat{x}_j \sim \left(\sum_{i=0}^{N_p} w^i_k\delta(x - x^i_k)\right)
\end{equation}

\begin{algorithm}
\caption{Resampling Algorithm}
\begin{algorithmic}
\STATE Calculate total weight, $W = \sum_{i=0}^{N_p} w^i$
\FORALL{$0 < i < N_p$}
    \STATE Draw $V$ from uniform range $[0, W]$
    \STATE $C = W_t$
    \FORALL{$0 < j < N_p$ and $C < V$}
        \STATE $C = C - w^j$
    \ENDFOR
    \STATE Add $[x^j, \frac{1}{N_p}]$ to the new distribution
\ENDFOR
\STATE 
\end{algorithmic}
\label{alg:Resampling}
\end{algorithm}
For infinite particles this new distribution will match the old.
Unfortunately, this isn't the truth in practice: since the support is
still limited to the original particles, the number of \emph{unique} particles can only go down.
This effect, dubbed \emph{particle impoverishment} can result in excessive quantization
errors in the final distribution. However, there is a solution. Instead of sampling from the
discrete distribution, a smoothing kernel is applied, and particles are drawn from
that distribution. Because it is continuous, particle impoverishment
cannot occur. The easiest way to sample from the continuous distribution is to break the 
re-sampling down into two steps. After calculating an estimate of the scale of the original
distribution, algorithm \autoref{alg:Resampling} is performed. Next, a distribution is generated
based on the variance of the original distributions.
Finally, for each particle in the discretely re-sampled distribution, a sample is drawn from 
the smoothing 
distribution and added to the particle.  The process regularization is defined as:

@@

The ultimate effect of this regularized resampling is a convergence similar to simulated annealing
or a genetic algorithm. Versions of $x$ that are fit (give good measurements) spawn more children 
nearby which allow for more accurate estimation near points of high likelihood. 
As the variance of the estimated
$x$'s decrease, the radius in which children are spawned also decreases. Eventually the radius
will approach the width of the underlying uncertainty, $\nu_x$ and $\nu_y$.
